Figures

### 2020-02-18 figures/01_geometry.Rmd
create grid of slope and distance comparisons, closes #29

# setwd("figures")
dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
                       stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class), 
                                  by = c("llid" = "lagoslakeid"))

# plot of true deepest versus visual centers
gg_scatter_state <- ggplot(data = st_drop_geometry(dt)) + 
  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = state)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  xlab("Distance to shore from deepest point") +
  ylab("Distance to shore from visual center")

gg_scatter_reservoir <- ggplot(data = st_drop_geometry(dt)) + 
  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  xlab("Distance to shore from deepest point") +
  ylab("Distance to shore from visual center")

cowplot::plot_grid(gg_scatter_state + 
                     theme(axis.title.x = element_text(color = "white")), 
                   gg_scatter_reservoir + 
                     labs(color = "") +
                     theme(axis.title.y = element_blank(), 
                           axis.text.y = element_blank()), 
                   rel_widths = c(1, 0.8))

gg_pntrng_deepest <- st_drop_geometry(dt) %>%
    dplyr::filter(!is.na(reservoir_class)) %>%
  group_by(reservoir_class) %>%
  summarize(dist_deepest = list(tibble::enframe(
    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = reservoir_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Distance from shore to deepest point (m)") + xlab("")

@@

```{r }
# compare inlake slope against buffer slope

dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
                       stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, 
                                  buffer100m_slope_mean), 
                                  by = c("llid" = "lagoslakeid"))

plot(dt$inlake_slope, dt$buffer100m_slope_mean/10, 
     xlim = c(0, 3), ylim = c(0, 3))
abline(0, 1)

# View(arrange(dt, buffer100m_slope_mean))

>>>>

# setwd("figures")
dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
                       stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, shape_class), 
                                  by = c("llid" = "lagoslakeid"))

# plot of true deepest versus visual centers
gg_scatter_state <- ggplot(data = st_drop_geometry(dt)) + 
  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = state)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  xlab("Deepest point distance (m)") +
  ylab("Visual center distance (m)")

gg_scatter_shape <- st_drop_geometry(dt) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex")))) %>%
  ggplot() + 
  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  labs(color = "") +
  xlab("Deepest point distance (m)") +
  ylab("Visual center distance (m)")

gg_scatter_reservoir <- ggplot(data = st_drop_geometry(dt)) + 
  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  xlab("Deepest point distance (m)") +
  ylab("Visual center distance (m)")
```

```{r gg_distance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3}
gg_distance <- cowplot::plot_grid(gg_scatter_shape + 
                     theme(axis.title.x = element_text(color = "white")), 
                   gg_scatter_reservoir + 
                     labs(color = "") +
                     theme(axis.title.y = element_blank(), 
                           axis.text.y = element_blank()), 
                   rel_widths = c(1, 0.8))
gg_distance

gg_pntrng_deepest <- st_drop_geometry(dt) %>%
    dplyr::filter(!is.na(reservoir_class)) %>%
  group_by(reservoir_class) %>%
  summarize(dist_deepest = list(tibble::enframe(
    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = reservoir_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Deepest point distance (m)") + xlab("")

@@

```{r slope-compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
# compare inlake slope, buffer slope, and online slope
# setwd("figures") 
bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(llid))
dt <- left_join(dt, st_drop_geometry(bp), 
                by = "llid") %>%
  dplyr::filter(!is.na(inlake_slope_mean))

# jsta::jheatmap(dt)
# jsta::jheatmap(dt, "inlake_slope.y")
jsta::jheatmap(
  dplyr::select(dt, 
                slope_online_mean,
                inlake_slope_mean,
                slope_mean, 
                slope_mean_ne,
                dist_deepest, 
                dist_viscenter, 
                maxdepth,
                meandepth))

GGally::ggpairs(dplyr::select(dt, 
                inlake_slope_mean, inlake_slope.y,
                slope_online_mean,
                slope_mean, 
                # dist_deepest, 
                # dist_viscenter, 
                maxdepth), 
                progress = FALSE, 
                ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
          strip.text.y = element_text(angle = 0))

gg_scatter_slope <- function(x, y, dt){
  ggplot(data = dt) + 
    geom_point(aes(UQ(rlang::sym(x)), UQ(rlang::sym(y))))
}

# dt[22,]
# which(dt$inlake_slope.y > 0.4)
# which(dt$inlake_slope.x > 0.4)
dt <- dplyr::filter(dt, inlake_slope.y < 0.4)

cowplot::plot_grid(
  gg_scatter_slope("slope_mean", "maxdepth", dt),
  gg_scatter_slope("inlake_slope.y", "maxdepth", dt), 
  gg_scatter_slope("slope_mean", "inlake_slope.y", dt)
  )
```

```{r buffer-vs-inlake_slope, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4}
lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex"))))

# GGally::ggpairs(dplyr::select(test, matches("class")))

gg_slope_shape <- ggplot(data = test, 
                         aes(slope_mean, inlake_slope.y, 
                             color = shape_class)) + 
    geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Buffer slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")

gg_slope_reservoir <- ggplot(data = test, 
                             aes(slope_mean, inlake_slope.y, 
                                 color = reservoir_class)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Buffer slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")
```

```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5}
theme_opts <- theme(legend.position = "none")
cowplot::plot_grid(gg_scatter_shape + 
                     labs(color = "") +
                     # theme(axis.title.x = element_blank()) +
                     theme_opts,
                   gg_scatter_reservoir + 
                     labs(color = "") +
                     theme(#axis.title.x = element_blank(), 
                           axis.title.y = element_blank()) +
                     theme_opts,
                   gg_slope_shape +
                     theme_opts,
                   gg_slope_reservoir + 
                     theme(axis.title.y = element_blank()) +
                     theme_opts, 
                   rel_widths = c(1, 0.93, 1, 1))
```

### 2020-02-18 manuscript/figures.Rmd
size adjustment
add slope comparison figure
create grid of slope and distance comparisons, closes #29

![Diagram showing the assumption of most lake depth models that the nearshore slope is extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope. ](../figures/slope_diagram.pdf){width=90%}

@@

```{r 01_geometry_points, fig.cap = "Distance to deepest point versus distance to lake visual center.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_geometry-1.pdf")
```

@@

```{r constrasts_distance, fig.cap = "Contrasting distance from shore to the deepest point in the lake and distance to the lake visual center for natural lakes and reservoirs.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_geometry-2.pdf")
```

@@

```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_geometry-3.pdf")
```

>>>>

![Diagram showing the assumption of most lake depth models that the nearshore slope is extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope. ](../figures/slope_diagram.pdf){width=85%}

@@

```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
```

@@

```{r constrasts_distance, fig.cap = "Contrasting distance from shore to the deepest point in the lake and distance to the lake visual center for natural lakes and reservoirs.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/gg_distance-2.pdf")
```

@@

```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/gg_distance-3.pdf")
```
